\subsection{Machine environment}

The experiments were executed across several AWS EC2 machines, as they offer
quite powerful virtual machines.

\subsubsection{Detailed description of machine environemnt}

The experiments ran concurrently in their own process with multithreading
turned on for the frameworks that supported it. The experiments ran on a
mixture of instances of type \texttt{c5.9xlarge} and
\texttt{c5.18xlarge}.\footnote{For the difference between instance types see
\url{https://aws.amazon.com/ec2/instance-types/}}
Each machine had a 32 GiB magnetic harddisk attached as its primary storage.
The key difference between the machines was the vCPU count, the memory
difference had no impact on execution.

Each machine ran software with the following versions:
\begin{itemize}
  \item{Ubuntu 18.04.2 LTS}
  \item{Python 3.6.7}
  \begin{itemize}
    \item{NumPy 1.16.2}
    \item{Keras 2.2.4 / TensorFlow 1.13.1}
    \item{dyNET 2.1}
    \item{PyTorch 1.0.1.post2}
  \end{itemize}
\end{itemize}

For DyNet 34 processes were executed concurrently depending on the
instance type.
For tensorflow 5 processes were executed concurrently.
Pytorch initally ran 2 processes concurrently with multithreading enabled, but
was changed to 34 processes with multithreading disabled midway through the
execution of the experiments.

\subsubsection{Differences between AWS and development environment}

Notable differences between execution in the development environment and on the
EC2 servers were noticed and are summarised by the following:

Pytorch grinded to a halt on the 36 core EC2 machine. Switching to a 72 core EC2
machine without disabling multithreading only exacerbated the problem. It turns
out that pytorch's multithreading does not support a high amount of
cores\footnote{As evidenced in this Github issue \url{https://github.com/pytorch/pytorch/issues/9873}}.

Tensorflow had no noticable differences between the development environment and
the EC2 machines.

Some DyNet configurations took more than six times the time to complete on an
EC2 machine compared to the local development environment. The exact cause is
not known to us.

\subsubsection{Comparability of speed results}

Given the complications stated in the above sections, none of the speed results
are truly comparable. This a result of several factors.
One being that 2 different EC2 instance types was used, where one was faster
than the other, or in pytorch's case slower.
Another is the fact that each experiment was run concurrently with others. This
would have only had a minor impact, if multithreading had not been enabled for
some of the frameworks.
Lastly DyNet behaved unexpectedly with regards to execution time, compared to the
development environment.

Ideally each experiment would be run one at a time, or with multithreading
disabled, on identical EC2 machines. Seeing as that was not the case, no
conclusions can be made with regards to the speed of the frameworks, only hints
at where further experimentation with tighter control might be warranted.
% Here we could add something about master/slave pattern og maybe how utilising
% AWS Batch would have been "oplagt", sorta, tror ogsaa det ville have vaeret
% lidt besvaerligt :/
